```{r}
chooseCRANmirror(graphics=FALSE, ind=1)
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyr)
library(ggplot2)
BPRS<- read.table("https://raw.githubusercontent.com/KimmoVehkalahti/MABS/master/Examples/data/BPRS.txt", sep  =" ", header = T)
BPRS$treatment <- factor(BPRS$treatment)
BPRS$subject <- factor(BPRS$subject)
BPRSL <-  BPRS %>% gather(key = weeks, value = bprs, -treatment, -subject)
BPRSL <-  BPRSL %>% mutate(week = as.integer(substr(weeks,5,5)))
ggplot(BPRSL, aes(x = week, y = bprs, linetype = subject)) +
  geom_line() +
  scale_linetype_manual(values = rep(1:10, times=4)) +
  facet_grid(. ~ treatment, labeller = label_both) +
  theme(legend.position = "none") + 
  scale_y_continuous(limits = c(min(BPRSL$bprs), max(BPRSL$bprs)))
```
The plot the BPRS score of almost all the men is decreasing over the eight weeks of the study.The men who have higher BPRS values at the beginning tend to have higher values throughout the study. There are substantial individual differences and variability appears to decrease with time. 

```{r}
BPRSL <- BPRSL %>%
  group_by(week) %>%
  mutate(stdbprs = (bprs - mean(bprs))/sd(bprs) ) %>%
  ungroup()
glimpse(BPRSL)
ggplot(BPRSL, aes(x = week, y = stdbprs, linetype = subject)) +
  geom_line() +
  scale_linetype_manual(values = rep(1:10, times=4)) +
  facet_grid(. ~ treatment, labeller = label_both) +
  scale_y_continuous(name = "standardized bprs")
```
The standarised plot shows more clearly the tracking phenomenon. 

```{r}
n <- BPRSL$week %>% unique() %>% length()
BPRSS <- BPRSL %>%
  group_by(treatment, week) %>%
  summarise( mean = mean(bprs), se = sd(bprs)/sqrt(n) ) %>%
  ungroup()
glimpse(BPRSS)
ggplot(BPRSS, aes(x = week, y = mean, linetype = treatment, shape = treatment)) +
  geom_line() +
  scale_linetype_manual(values = c(1,2)) +
  geom_point(size=3) +
  scale_shape_manual(values = c(1,2)) +
  geom_errorbar(aes(ymin = mean - se, ymax = mean + se, linetype="1"), width=0.3) +
  theme(legend.position = c(0.8,0.8)) +
  scale_y_continuous(name = "mean(bprs) +/- se(bprs)")
```
The summary graph displays that both treatment groups' BPRS values are decreasing in general. Treatment 2's BPRS values however, rebounce after five weeks. The speed of decreasing BPRS values slows down in treament 1 after 6 weeks. 

```{r}
BPRSL8S <- BPRSL %>%
  filter(week > 0) %>%
  group_by(treatment, subject) %>%
  summarise( mean=mean(bprs) ) %>%
  ungroup()
glimpse(BPRSL8S)
ggplot(BPRSL8S, aes(x = treatment, y = mean)) +
  geom_boxplot() +
  stat_summary(fun.y = "mean", geom = "point", shape=23, size=4, fill = "white") +
  scale_y_continuous(name = "mean(bprs), weeks 1-8")
```
The diagram indicates that the mean summary measure is more variable in the second treatment group and its distribution in this group is somewhat skew. The boxplot of the second group reveals an outlier,
a subject whose mean BPRS score of the eight weeks is over 70. 


```{r}
BPRSL8S1 <- BPRSL8S %>%
  filter(mean < 60) %>%
  group_by(treatment, subject) %>%
  summarise(mean=mean(mean)) %>%
  ungroup()
glimpse(BPRSL8S1)
ggplot(BPRSL8S1, aes(x = treatment, y = mean)) +
  geom_boxplot() +
  stat_summary(fun.y = "mean", geom = "point", shape=23, size=4, fill = "white") +
  scale_y_continuous(name = "mean(bprs), weeks 1-8")

```
After remove the outliner, the mean of the treatment 2 becomes smaller than treatment 1. We still cannot tell if the two treatment has significant difference. 

```{r}
t.test(mean ~ treatment, data = BPRSL8S1, var.equal = TRUE)
BPRSL8S2 <- BPRSL8S %>%
  mutate(baseline = BPRS$week0)
fit <- lm(mean ~ baseline + treatment, data = BPRSL8S2)
anova(fit)
```
The t-test confirms the lack of any evidence for a group difference. Also the
95% confidence interval is wide and includes the zero, allowing for similar
conclusions to be made.

####Part2
```{r}
RATS <- read.table("https://raw.githubusercontent.com/KimmoVehkalahti/MABS/master/Examples/data/rats.txt", sep = "", header = T)
RATS$Group <- factor(RATS$Group)
RATS$ID <- factor(RATS$ID)
RATSL <-  RATS %>% gather(key = WDs, value = Weight, -Group, -ID)
RATSL <-  RATSL %>% mutate(Time = as.integer(substr(WDs,3,4)))
ggplot(RATSL, aes(x = Time, y = Weight,group = ID)) +
  geom_line(aes(linetype = Group)) +
  scale_x_continuous(name = "Time (days)", breaks = seq(0, 60, 10)) +
  scale_y_continuous(name = "Weight (grams)") +
  theme(legend.position = "top")
```
This plot plots time and weight of the rats as three groups. Group1 seems to be clearly different (smaller weight) from other two groups.

```{r}
RATS_reg <- lm(Weight ~ Time + Group, data = RATSL)
summary(RATS_reg)
```
```{r}
library(lme4)
RATS_ref <- lmer(Weight ~ Time + Group + (1 | ID), data = RATSL, REML = FALSE)
summary(RATS_ref)
```

The estimated variance of the rat random effects is quite large, indicating the considerable variation in the intercepts of the regression fits of the individual rat growth profiles.
```{r}
RATS_ref1 <- lmer(Weight ~ Time + Group + (Time | ID), data = RATSL, REML = FALSE)
summary(RATS_ref1)
anova(RATS_ref1, RATS_ref)
```
The likelihood ratio test of the interaction random intercept and slope model against the corresponding
model without an interaction is 12.36 with 2 DF; the associated p-value is very small, and we can conclude that the interaction model provides a better fit for the rat growth data. The estimated regression parameters for the interaction in Table 9.6 indicate that the growth rate slopes are considerably higher for rats in group 2 than for rats in group 1 (on average 0.61 higher with an approximate 95% confidence interval [CI] of [0.33, 0.89]) but less so when comparing group 3 rats with those in group 1 (on average 0.30 higher, CI
[0.02, 0.58]).

```{r}
RATS_ref2 <- lmer(Weight ~ Time * Group + (Time | ID), data = RATSL, REML = FALSE)
summary(RATS_ref2)
anova(RATS_ref2, RATS_ref1)
ggplot(RATSL, aes(x = Time, y = Weight, group = ID)) +
  geom_line(aes(linetype = Group)) +
  scale_x_continuous(name = "Time (days)", breaks = seq(0, 60, 20)) +
  scale_y_continuous(name = "Observed weight (grams)") +
  theme(legend.position = "top")
Fitted <- fitted(RATS_ref2)
RATSL <- RATSL %>%
  mutate(Fitted)
ggplot(RATSL, aes(x = Time, y = Fitted, group = ID)) +
  geom_line(aes(linetype = Group)) +
  scale_x_continuous(name = "Time (days)", breaks = seq(0, 60, 20)) +
  scale_y_continuous(name = "Fitted weight (grams)") +
  theme(legend.position = "top")
```




